Improved Bounds for Online Balanced Repartitioning

### 2020-04-22 paper/disc20clustering.tex
On cascade
Generic algorithm

We assume \OPT begins with the initial configuration
$P_I = I_1, \dots, I_{\ell}$ and moves to the final partitioning
$P_F = F_1, \dots, F_{\ell}$.
 The \emph{distance} of a configuration $P = C_1, \dots, C_{\ell}$ from the initial configuration \maciek{, denoted $\dist(P)$} is the number of nodes in $P$ that do not reside in their initial cluster.
 \maciek{It is OK to use single-argument $\dist$ (two-arg is a must in a definition, and it is already there).}
 \maciek{We should denote distance $\Delta(P)$, and the current $\Delta$ should be R (like radius).}
    That is,
    $\dist(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$. 
In other words,
at least $\dist(P, P_I)/2$ node swaps are required in order to reach the configuration $P$ from $P_I$, and thus
$\OPT \geq \Delta:= \dist(P_F, P_I) $.
 With each repartitioning,
  \PPL moves to a configuration that minimizes the distance to the initial configuration $P_I$.
As a result,
\PPL never ends up in a configuration that is more than $\Delta$ (single-node) migrations away from $P_I$.
This invariant ensures that \PPL does not pay too much while recovering $P_F$.
%      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial partition}
      We note that the $\mathit{repartition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a~perfect partition closest to $P_I$.
Hence it never moves to a configuration beyond distance $\Delta$.
The scheme of the algorithm can be found in the Algorithm \ref{alg:PPL}.

@@

\begin{lemma}	\label{lemma:rebalancecost}
    The cost of repartitioning at Line \ref{line:rebalance} is at most $2\cdot\OPT$.
\end{lemma}
\begin{proof}
    Consider the repartitioning that transforms $P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
    Let $M \subset V$ denote the set of nodes that migrate during this process.
	Let $M^-$ and $M^+$ denote the subset of nodes that (respectively)
    enter or leave their original cluster during the repartitioning.    
    Then,
    $M = M^+ \cup M^-$.
    Since $|M^-|$ nodes are not in their original cluster before the repartitioning (i.e., in $P_{t-1}$),
    the distance before the repartitioning is $\dist(P_{t-1},P_I) \geq | M^-|$.
    Analogously,
     the distance afterwards is $\dist(P_{t},P_I) \geq | M^+|$.
    Thus,
    $|M| \leq \dist(P_{t-1},P_I) + \dist(P_{t},P_I)$.
    By Property \ref{prop:dist<OPT},
    $\dist(P_{t-1},P_I) , \dist(P_{t},P_I) \leq \Delta \leq \OPT$
    and thereby we have	
    $|M| \leq 2\cdot\OPT$.
    \maciek{the comma between $\dist$ is confusing. BOTH are bounded?}
\end{proof}

@@

\maciek{Better name than DET?}

@@

\begin{theorem}
  \TAlg is $O(\ell)$-competitive.
\end{theorem}
\begin{proof}
  Fix a completed phase, and consider the state of \TAlg's counters at the end of it.
  We consider the incomplete phase later in this proof.

@@

  Now we estimate the cost of $\OPT(\sigma_E)$ while accounting savings from migrations.
  By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
  This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
  As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
  By performing a single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$.
  Thus, for serving $\sigma_E$, $\OPT$ pays at least $\OPT(\sigma_E) + \OPT(M) \geq W \cdot \frac{2\alpha}{4 (\alpha-1)}\geq |\sigma_E| / 2 - 2 \alpha \cdot \ell$.
  From that, we obtain $|\sigma_E| \geq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
  Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, we are ready to bound the competitive ratio
  \begin{equation*}
    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + E/2} \leq \frac{27\alpha\cdot\ell + 2\cdot E}{\alpha + E} \leq 27 \ell = O(\ell).
  \end{equation*}

@@

  \maciek{Big techincal TODO. Formalize the sketched proof for the last phase.}
  It remains to consider the last, unfinished phase.
  Consider the case, where the unfinished phase is also the first one.
  Then, we cannot charge $\OPT$ due to the inability to partition the components.
  Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
  We charge $\OPT$ $\alpha$ for the first external $\alpha$ requests or a migration,
  and we follow the analysis regarding the unsaturated requests.
  If the input finished before the first $\alpha$ external requests, then \TAlg is optimal.
  Now, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
  This way, the competitive ratio increases at most twofold.
\end{proof}

\section{Bounding the cost of a single reconfiguration}

@@

\begin{proof}
  Each node has degree at most $k$. Combine with Theorem \ref{th:cascade-paths}.

>>>>

We assume \OPT begins with the initial configuration
$P_I = I_1, \dots, I_{\ell}$ and moves to the final partitioning
$P_F = F_1, \dots, F_{\ell}$.
 The \emph{distance} of a configuration $P = C_1, \dots, C_{\ell}$ from the initial configuration \maciek{, denoted $\dist(P)$} is the number of nodes in $P$ that do not reside in their initial cluster.
 \maciek{We might denote distance $\Delta(P)$, and the current $\Delta$ might be R (like radius).}
    That is,
    $\dist(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$. 
In other words,
at least $\dist(P, P_I)/2$ node swaps are required in order to reach the configuration $P$ from $P_I$, and thus
$\OPT \geq \Delta:= \dist(P_F, P_I) $.
 With each repartitioning,
  \PPL moves to a configuration that minimizes the distance to the initial configuration $P_I$.
As a result,
\PPL never ends up in a configuration that is more than $\Delta$ (single-node) migrations away from $P_I$.
This invariant ensures that \PPL does not pay too much while recovering $P_F$.
%      \maciek{``re-partition'' procedure name should indicate the fact that this is a specific repartition that is close to initial partition}
      We note that the $\mathit{repartition}$ at Line \ref{line:rebalance} replaces the current configuration $P$ with a~perfect partition closest to $P_I$.
Hence it never moves to a configuration beyond distance $\Delta$.
The scheme of the algorithm can be found in the Algorithm \ref{alg:PPL}.

@@

\begin{lemma}	\label{lemma:rebalancecost}
    The cost of repartitioning at Line \ref{line:rebalance} is at most $2\cdot\OPT$.
\end{lemma}
\begin{proof}
    Consider the repartitioning that transforms $P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
    Let $M \subset V$ denote the set of nodes that migrate during this process.
	Let $M^-$ and $M^+$ denote the subset of nodes that (respectively)
    enter or leave their original cluster during the repartitioning.    
    Then,
    $M = M^+ \cup M^-$.
    Since $|M^-|$ nodes are not in their original cluster before the repartitioning (i.e., in $P_{t-1}$),
    the distance before the repartitioning is $\dist(P_{t-1},P_I) \geq | M^-|$.
    Analogously,
     the distance afterwards is $\dist(P_{t},P_I) \geq | M^+|$.
    Thus,
    $|M| \leq \dist(P_{t-1},P_I) + \dist(P_{t},P_I)$.
    By Property \ref{prop:dist<OPT},
    $\dist(P_{t-1},P_I) , \dist(P_{t},P_I) \leq \Delta \leq \OPT$
    and thereby we have	
    $|M| \leq 2\cdot\OPT$.
    \maciek{the comma between $\dist$ is confusing. BOTH are bounded? If so, expand the sentence and write both individual bounds.}
\end{proof}

@@

\maciek{Better name than DET? Currently it is \TAlg, hence it is wrong}

@@

\begin{theorem}
  \TAlg is $O(\ell)$-competitive.
  \maciek{Let Reconf(k,l) be a bound on the cost of the reconfiguration. Then \TAlg is $O(Reconf(k,l)\cdot k \cdot \ell)$-competitive.}
\end{theorem}
\begin{proof}
  Fix a completed phase, and consider the state of \TAlg's counters at the end of it.
  We consider the incomplete phase later in this proof.

@@

  Now we estimate the cost of $\OPT(\sigma_E)$ while accounting savings from migrations.
  By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
  This may allow serving requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
  As $\sigma_E$ consists of requests to unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
  By performing a single swap that costs $2\alpha$, $\OPT$ may avoid paying the remote serving costs for at most $4 (\alpha - 1)$ requests from $\sigma_E$
  \maciek{This changes in the general case, and introduces the factor $~1/k$ in opt's cost, increasing the competitive ratio by $O(k)$}.
  Thus, for serving $\sigma_E$, $\OPT$ pays at least $\OPT(\sigma_E) + \OPT(M) \geq W \cdot \frac{2\alpha}{4 (\alpha-1)}\geq |\sigma_E| / 2 - 2 \alpha \cdot \ell$.
  From that, we obtain $|\sigma_E| \geq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
  Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, we are ready to bound the competitive ratio
  \begin{equation*}
    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + E/2} \leq \frac{27\alpha\cdot\ell + 2\cdot E}{\alpha + E} \leq 27 \ell = O(\ell).
  \end{equation*}

@@

  It remains to consider the last, unfinished phase.
  Consider the case, where the unfinished phase is also the first one.
  Then, we cannot charge $\OPT$ due to the inability to partition the components.
  Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
  We charge $\OPT$ $\alpha$ for the first external $\alpha$ requests or a migration,
  and we follow the analysis regarding the unsaturated requests.
  If the input finished before the first $\alpha$ external requests, then \TAlg is optimal.
  \maciek{Missing second case: then we just charge opt alpha for this one edge and this is sufficient, as the cost of alg only decreases from the full phase}
  Now, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
  This way, the competitive ratio increases at most twofold.
\end{proof}

\section{Bounding the cost of a single reconfiguration of components}

\maciek{Overview of the section, intro. Defer introducing the graph notation, the warm-up does not need it}

@@

\begin{proof}
  Each node has degree at most $k$.
  Every subset has flow conservation property, hence each path traverses one of the nodes of graph origin/core. (no subsets with just incoming edges to the origin part)
  Combine with Theorem \ref{th:cascade-paths}.
